---
title: "Gene Set Enrichment analysis part III: snRNAseq samples with metformin induced cardiac microtissue"
subtitle: "Adapted from Gino Bonazza DCM samples analysis by Magdalena Czepiec and Maks Necki"
author:
  - name: |
      <b>Magdalena Czepiec</b><br>
      Department of Clinical Immunology / Jagiellonian University<br>
      <a href="https://github.com/czepiecmagdalena">GitHub</a>

  - name: |
      <b>Maks Necki</b><br>
      Department of Clinical Immunology / Jagiellonian University<br>
      <a href="https://github.com/jungwirt">GitHub</a>

tags: [snRNAseq, heart microtissue, metformin, GSEA, category scores]

fontsize: 11pt
geometry: margin=1in
lang: en

output:
  rmdformats::downcute:
    thumbnails: true
    lightbox: true
    gallery: true
    highlight: tango
    number_sections: true
    code_folding: hide
    df_print: paged
    self_contained: true

editor_options:
  chunk_output_type: console

params:
  organism: "Homo sapiens"
  analysis_type: "GSEA"
---
```{r setup, include=FALSE}
tags <- rmarkdown::metadata$tags
```

# Introduction
<div style="text-align: justify;">
A module score represents a quantitative metric that summarizes the coordinated expression of a predefined gene set (e.g., genes associated with metabolism, stress response, or signaling pathways) into a single value for each cell or sample.
It is computed by aggregating the expression levels of genes within a given biological category—often with normalization against control gene sets—thereby providing an integrated measure of pathway or process activity rather than relying on individual gene-level signals.
This approach enables:
- comparative assessment of pathway activity across cells, experimental conditions, or clusters,
- identification of cellular subpopulations characterized by differential activation of specific biological processes,
- reduction of data dimensionality, facilitating the interpretation of complex transcriptomic datasets.
</div>

## Setup: Libraries and Environement

```{r installing and loading libraries, message=FALSE, warning=FALSE}
  # ---- Package lists ----
  cran_packages <- c(
  "dplyr",
  "ggplot2",
  "patchwork",
  "htmltools",
  "tidyr",
  "Seurat",
  "stringr",
  "plotly",
  "kableExtra",
  "knitr",
  "sessioninfo"
  )

  bioc_packages <- c(
  "clusterProfiler",
  "org.Hs.eg.db",
  "enrichplot",
  "ReactomePA"
  )

  # ---- Function to install & load ----
  install_and_load <- function(pkg, bioc = FALSE) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    if (bioc) {
      if (!requireNamespace("BiocManager", quietly = TRUE)) {
        install.packages("BiocManager")
      }
      BiocManager::install(pkg, ask = FALSE, update = FALSE)
    } else {
      install.packages(pkg)
    }
  }
  library(pkg, character.only = TRUE)
  }

  # ---- Install & load CRAN packages ----
  invisible(lapply(cran_packages, install_and_load))

  # ---- Install & load Bioconductor packages ----
  invisible(lapply(bioc_packages, install_and_load, bioc = TRUE))

  cat("ALL PACKAGES SUCCESSFULLY LOADED")
```

## Helper function for importing extraction datasets without predefined gene lists (no genes after GSEA part II)
```{r}
  safe_read_csv <- function(path) {
  if (!file.exists(path)) {
    message("No file found (likely no genes / empty result): ", path)
    return(NULL)
  }

  df <- tryCatch(
    read.csv(path),
    error = function(e) NULL
  )

  if (is.null(df) || nrow(df) == 0) {
    message("File is empty or invalid: ", path)
    return(NULL)
  }

  return(df)
  }
```

## Importing files after extraction (GSEA part II)

```{r}
# ---- Metabolism ----
score_mettgf_tgf_metabolism <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_metabolism/mettgf_vs_tgf_Metabolism.csv")
score_tgf_nc_metabolism <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_metabolism/tgf_vs_nc_Metabolism.csv")
score_met_nc_metabolism <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_metabolism/met_vs_nc_Metabolism.csv")

# ---- Immune ----
score_mettgf_tgf_immune <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_immune/mettgf_vs_tgf_Immune.csv")
score_tgf_nc_immune <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_immune/tgf_vs_nc_Immune.csv")
score_met_nc_immune <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_immune/met_vs_nc_Immune.csv")

# ---- Migration ----
score_mettgf_tgf_migration <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_migration/mettgf_vs_tgf_Migration.csv")
score_tgf_nc_migration <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_migration/tgf_vs_nc_Migration.csv")
score_met_nc_migration <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_migration/met_vs_nc_Migration.csv")

# ---- Cell cycle ----
score_mettgf_tgf_cellcycle <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_cellcycle/mettgf_vs_tgf_Cell_cycle.csv")
score_tgf_nc_cellcycle <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_cellcycle/tgf_vs_nc_Cell_cycle.csv")
score_met_nc_cellcycle <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_cellcycle/met_vs_nc_Cell_cycle.csv")

# ---- Signaling ----
score_mettgf_tgf_signaling <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_signaling/mettgf_vs_tgf_Signaling.csv")
score_tgf_nc_signaling <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_signaling/tgf_vs_nc_Signaling.csv")
score_met_nc_signaling <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_signaling/met_vs_nc_Signaling.csv")

# ---- Development ----
score_mettgf_tgf_development <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_development/mettgf_vs_tgf_Development.csv")
score_tgf_nc_development <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_development/tgf_vs_nc_Development.csv")
score_met_nc_development <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_development/met_vs_nc_Development.csv")

# ---- Cell death ----
score_mettgf_tgf_celldeath <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_celldeath/mettgf_vs_tgf_Cell_death.csv")
score_tgf_nc_celldeath <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_celldeath/tgf_vs_nc_Cell_death.csv")
score_met_nc_celldeath <- safe_read_csv("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/results_fibro_extraction_celldeath/met_vs_nc_Cell_death.csv")
```

## Importing Seurat object containing analyzed and processed fibroblast cells from all conditions

```{r}
Fibro <- readRDS("C:/Users/magda/Desktop/Lib_11_MTs_GSEA/Skrypty do publikacji/GSEA part II (Gene Extraction)/Fibroblasts/Fibro.rds")
```

# Computing module scores for each biological category

```{r}
  # ---- Genes within metabolic category ----
  genes_tgf_nc_metabolism <- unique(as.character(score_tgf_nc_metabolism$x))
  genes_mettgf_tgf_metabolism <- unique(as.character(score_mettgf_tgf_metabolism$x))
  genes_met_nc_metabolism <- unique(as.character(score_met_nc_metabolism$x))

  genes_all_metabolism <- unique(c(
  genes_met_nc_metabolism,
  genes_mettgf_tgf_metabolism,
  genes_tgf_nc_metabolism
  ))

  genes_intersect_list_metabolism <- intersect(genes_all_metabolism, rownames(Fibro))

  if (length(genes_intersect_list_metabolism) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_metabolism))
  
  Fibro$Metabolism_Score_all1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_metabolism),
    name = "Metabolism_Score_all"
  )
  }

# ---- Genes within immune category ----
  genes_tgf_nc_immune <- unique(as.character(score_tgf_nc_immune$x))
  genes_mettgf_tgf_immune <- unique(as.character(score_mettgf_tgf_immune$x))
  genes_met_nc_immune <- unique(as.character(score_met_nc_immune$x))

  genes_all_immune <- unique(c(
  genes_met_nc_immune,
  genes_mettgf_tgf_immune,
  genes_tgf_nc_immune
  ))

  genes_intersect_list_immune <- intersect(genes_all_immune, rownames(Fibro))

  if (length(genes_intersect_list_immune) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_immune))
  
  Fibro$Immune_Score_all1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_immune),
    name = "Immune_Score_all"
  )
  }

  # ---- Genes within migration category ----
  genes_tgf_nc_migration <- unique(as.character(score_tgf_nc_migration$x))
  genes_mettgf_tgf_migration <- unique(as.character(score_mettgf_tgf_migration$x))
  genes_met_nc_migration <- unique(as.character(score_met_nc_migration$x))

  genes_all_migration <- unique(c(
  genes_met_nc_migration,
  genes_mettgf_tgf_migration,
  genes_tgf_nc_migration
  ))

  genes_intersect_list_migration <- intersect(genes_all_migration, rownames(Fibro))

  if (length(genes_intersect_list_migration) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_migration))
  
  Fibro$Migration_Score_all1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_migration),
    name = "Migration_Score_all"
  )
  }

  # ---- Genes within cell cycle category ----
  genes_tgf_nc_cellcycle <- unique(as.character(score_tgf_nc_cellcycle$x))
  genes_mettgf_tgf_cellcycle <- unique(as.character(score_mettgf_tgf_cellcycle$x))
  genes_met_nc_cellcycle <- unique(as.character(score_met_nc_cellcycle$x))

  genes_all_cellcycle <- unique(c(
  genes_met_nc_cellcycle,
  genes_mettgf_tgf_cellcycle,
  genes_tgf_nc_cellcycle
  ))

  genes_intersect_list_cellcycle <- intersect(genes_all_cellcycle, rownames(Fibro))

  if (length(genes_intersect_list_cellcycle) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_cellcycle))
  
  Fibro$Cell_cycle_Score_all1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_cellcycle),
    name = "Cell_cycle_Score_all"
  )
  }

  # ---- Genes within signaling category ----
  genes_tgf_nc_signaling <- unique(as.character(score_tgf_nc_signaling$x))
  genes_mettgf_tgf_signaling <- unique(as.character(score_mettgf_tgf_signaling$x))
  genes_met_nc_signaling <- unique(as.character(score_met_nc_signaling$x))

  genes_all_signaling <- unique(c(
  genes_met_nc_signaling,
  genes_mettgf_tgf_signaling,
  genes_tgf_nc_signaling
  ))

  genes_intersect_list_signaling <- intersect(genes_all_signaling, rownames(Fibro))

  if (length(genes_intersect_list_signaling) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_signaling))
  
  Fibro$Signaling_Score_all1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_signaling),
    name = "Signaling_Score_all"
  )
  }

  # ---- Genes within development category ----
  genes_tgf_nc_development <- unique(as.character(score_tgf_nc_development$x))
  genes_mettgf_tgf_development <- unique(as.character(score_mettgf_tgf_development$x))
  genes_met_nc_development <- unique(as.character(score_met_nc_development$x))

  genes_all_development <- unique(c(
  genes_met_nc_development,
  genes_mettgf_tgf_development,
  genes_tgf_nc_development
  ))

  genes_intersect_list_development <- intersect(genes_all_development, rownames(Fibro))

  if (length(genes_intersect_list_development) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_development))
  
  Fibro$Development_Score_all1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_development),
    name = "Development_Score_all"
  )
  }

  # ---- Genes within cell death category ----
  genes_tgf_nc_celldeath <- unique(as.character(score_tgf_nc_celldeath$x))
  genes_mettgf_tgf_celldeath <- unique(as.character(score_mettgf_tgf_celldeath$x))
  genes_met_nc_celldeath <- unique(as.character(score_met_nc_celldeath$x))

  genes_all_celldeath <- unique(c(
  genes_met_nc_celldeath,
  genes_mettgf_tgf_celldeath,
  genes_tgf_nc_celldeath
  ))

  genes_intersect_list_celldeath <- intersect(genes_all_celldeath, rownames(Fibro))

  if (length(genes_intersect_list_celldeath) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_celldeath))
  
  Fibro$Cell_death_Score_all1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_celldeath),
    name = "Cell_death_Score_all"
  )
  }
```

## Visualizing Metabolic Scores

```{r}
VlnPlot(Fibro, features = "Metabolism_Score_all1", group.by = "condition")

p <- VlnPlot(Fibro, features = "Metabolism_Score_all1", group.by = "condition")
ggsave(
  filename = "Metabolism_Score_all1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Migration Scores

```{r}
VlnPlot(Fibro, features = "Migration_Score_all1", group.by = "condition")

p <- VlnPlot(Fibro, features = "Migration_Score_all1", group.by = "condition")
ggsave(
  filename = "Migration_Score_all1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Immune Scores

```{r}
VlnPlot(Fibro, features = "Immune_Score_all1", group.by = "condition")

p <- VlnPlot(Fibro, features = "Immune_Score_all1", group.by = "condition")
ggsave(
  filename = "Immune_Score_all1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Cell Cycle Scores

```{r}
VlnPlot(Fibro, features = "Cell_cycle_Score_all1", group.by = "condition")

p <- VlnPlot(Fibro, features = "Cell_cycle_Score_all1", group.by = "condition")
ggsave(
  filename = "Cell_cycle_Score_all1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Signaling Scores

```{r}
VlnPlot(Fibro, features = "Signaling_Score_all1", group.by = "condition")

p <- VlnPlot(Fibro, features = "Signaling_Score_all1", group.by = "condition")
ggsave(
  filename = "Signaling_Score_all1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Development Scores

```{r}
VlnPlot(Fibro, features = "Development_Score_all1", group.by = "condition")

p <- VlnPlot(Fibro, features = "Development_Score_all1", group.by = "condition")
ggsave(
  filename = "Development_Score_all1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Cell Death Scores

```{r}
VlnPlot(Fibro, features = "Cell_death_Score_all1", group.by = "condition")

p <- VlnPlot(Fibro, features = "Cell_death_Score_all1", group.by = "condition")
ggsave(
  filename = "Cell_death_Score_all1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```


# Computing module scores for each biological category only for MET_TGF vs TGF comparison

```{r}
  # ---- Genes within metabolic category ----
  genes_mettgf_tgf_metabolism <- unique(as.character(score_mettgf_tgf_metabolism$x))

  genes_intersect_list_metabolism_mettgf_tgf <- intersect(genes_mettgf_tgf_metabolism, rownames(Fibro))

  if (length(genes_intersect_list_metabolism_mettgf_tgf) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_metabolism_mettgf_tgf))
  
  Fibro$Metabolism_Score_mettgf_tgf1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_metabolism_mettgf_tgf),
    name = "Metabolism_Score_mettgf_tgf"
  )
  }

# ---- Genes within immune category ----
  genes_mettgf_tgf_immune <- unique(as.character(score_mettgf_tgf_immune$x))

  genes_intersect_list_immune_mettgf_tgf <- intersect(genes_mettgf_tgf_immune, rownames(Fibro))

  if (length(genes_intersect_list_immune_mettgf_tgf) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_immune_mettgf_tgf))
  
  Fibro$Immune_Score_mettgf_tgf1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_immune_mettgf_tgf),
    name = "Immune_Score_mettgf_tgf"
  )
  }

  # ---- Genes within migration category ----
  genes_mettgf_tgf_migration <- unique(as.character(score_mettgf_tgf_migration$x))

  genes_intersect_list_migration_mettgf_tgf <- intersect(genes_mettgf_tgf_migration, rownames(Fibro))

  if (length(genes_intersect_list_migration_mettgf_tgf) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_migration_mettgf_tgf))
  
  Fibro$Migration_Score_mettgf_tgf1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_migration_mettgf_tgf),
    name = "Migration_Score_mettgf_tgf"
  )
  }

  # ---- Genes within cell cycle category ----
  genes_mettgf_tgf_cellcycle <- unique(as.character(score_mettgf_tgf_cellcycle$x))

  genes_intersect_list_cellcycle_mettgf_tgf <- intersect(genes_mettgf_tgf_cellcycle, rownames(Fibro))

  if (length(genes_intersect_list_cellcycle_mettgf_tgf) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_cellcycle_mettgf_tgf))
  
  Fibro$Cell_cycle_Score_mettgf_tgf1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_cellcycle_mettgf_tgf),
    name = "Cell_cycle_Score_mettgf_tgf"
  )
  }

  # ---- Genes within signaling category ----
  genes_mettgf_tgf_signaling <- unique(as.character(score_mettgf_tgf_signaling$x))

  genes_intersect_list_signaling_mettgf_tgf <- intersect(genes_mettgf_tgf_signaling, rownames(Fibro))

  if (length(genes_intersect_list_signaling_mettgf_tgf) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_signaling_mettgf_tgf))
  
  Fibro$Signaling_Score_mettgf_tgf1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_signaling_mettgf_tgf),
    name = "Signaling_Score_mettgf_tgf"
  )
  }

  # ---- Genes within development category ----
  genes_mettgf_tgf_development <- unique(as.character(score_mettgf_tgf_development$x))

  genes_intersect_list_development_mettgf_tgf <- intersect(genes_mettgf_tgf_development, rownames(Fibro))

  if (length(genes_intersect_list_development_mettgf_tgf) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_development_mettgf_tgf))
  
  Fibro$Development_Score_mettgf_tgf1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_development_mettgf_tgf),
    name = "Development_Score_mettgf_tgf"
  )
  }

  # ---- Genes within cell death category ----
  genes_mettgf_tgf_celldeath <- unique(as.character(score_mettgf_tgf_celldeath$x))

  genes_intersect_list_celldeath_mettgf_tgf <- intersect(genes_mettgf_tgf_celldeath, rownames(Fibro))

  if (length(genes_intersect_list_celldeath_mettgf_tgf) < 5) {
  message("Too few genes for module score: ", length(genes_intersect_list_celldeath_mettgf_tgf))
  
  Fibro$Cell_death_Score_mettgf_tgf1 <- NA
  
  } else {
  
  Fibro <- AddModuleScore(
    Fibro,
    features = list(genes_intersect_list_celldeath_mettgf_tgf),
    name = "Cell_death_Score_mettgf_tgf"
  )
  }
```

## Visualizing Metabolic Scores for mettgf_tgf comparision only

```{r}
VlnPlot(Fibro, features = "Metabolism_Score_mettgf_tgf1", group.by = "condition")

p1 <- VlnPlot(Fibro, features = "Metabolism_Score_mettgf_tgf1", group.by = "condition")
ggsave(
  filename = "Metabolism_Score_mettgf_tgf1.jpg",
  plot = p1,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Migration Scores for mettgf_tgf comparision only

```{r}
VlnPlot(Fibro, features = "Migration_Score_mettgf_tgf1", group.by = "condition")

p1 <- VlnPlot(Fibro, features = "Migration_Score_mettgf_tgf1", group.by = "condition")
ggsave(
  filename = "Migration_Score_mettgf_tgf1.jpg",
  plot = p1,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Immune Scores for mettgf_tgf comparision only

```{r}
VlnPlot(Fibro, features = "Immune_Score_mettgf_tgf1", group.by = "condition")

p1 <- VlnPlot(Fibro, features = "Immune_Score_mettgf_tgf1", group.by = "condition")
ggsave(
  filename = "Immune_Score_mettgf_tgf1.jpg",
  plot = p1,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Cell Cycle Scores for mettgf_tgf comparision only

```{r}
VlnPlot(Fibro, features = "Cell_cycle_Score_mettgf_tgf1", group.by = "condition")

p1 <- VlnPlot(Fibro, features = "Cell_cycle_Score_mettgf_tgf1", group.by = "condition")
ggsave(
  filename = "Cell_cycle_Score_mettgf_tgf1.jpg",
  plot = p1,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Signaling Scores for mettgf_tgf comparision only

```{r}
VlnPlot(Fibro, features = "Signaling_Score_mettgf_tgf1", group.by = "condition")

p1 <- VlnPlot(Fibro, features = "Signaling_Score_mettgf_tgf1", group.by = "condition")
ggsave(
  filename = "Signaling_Score_mettgf_tgf1.jpg",
  plot = p1,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Development Scores for mettgf_tgf comparision only

```{r}
VlnPlot(Fibro, features = "Development_Score_mettgf_tgf1", group.by = "condition")

p1 <- VlnPlot(Fibro, features = "Development_Score_mettgf_tgf1", group.by = "condition")
ggsave(
  filename = "Development_Score_mettgf_tgf1.jpg",
  plot = p1,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Cell Death Scores for mettgf_tgf comparision only

```{r}
VlnPlot(Fibro, features = "Cell_death_Score_mettgf_tgf1", group.by = "condition")

p1 <- VlnPlot(Fibro, features = "Cell_death_Score_mettgf_tgf1", group.by = "condition")
ggsave(
  filename = "Cell_death_Score_mettgf_tgf1.jpg",
  plot = p1,
  width = 8,
  height = 6,
  dpi = 300
)
```

# Computing module scores: compare all vs MET_TGF vs TGF comparison only

## Visualizing Metabolic Scores compare all vs mettgf_tgf
```{r}
p <- VlnPlot(
  Fibro,
  features = c("Metabolism_Score_all1", "Metabolism_Score_mettgf_tgf1"),
  group.by = "condition",
  combine = TRUE
)
p
ggsave(
  filename = "Metabolism_Score_combine1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Migration Scores compare all vs mettgf_tgf
```{r}
p <- VlnPlot(
  Fibro,
  features = c("Migration_Score_all1", "Migration_Score_mettgf_tgf1"),
  group.by = "condition",
  combine = TRUE
)
p
ggsave(
  filename = "Migration_Score_combine1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Immune Scores compare all vs mettgf_tgf
```{r}
p <- VlnPlot(
  Fibro,
  features = c("Immune_Score_all1", "Immune_Score_mettgf_tgf1"),
  group.by = "condition",
  combine = TRUE
)
p
ggsave(
  filename = "Immune_Score_combine1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Cell Cycle Scores compare all vs mettgf_tgf
```{r}
p <- VlnPlot(
  Fibro,
  features = c("Cell_cycle_Score_all1", "Cell_cycle_Score_mettgf_tgf1"),
  group.by = "condition",
  combine = TRUE
)
p
ggsave(
  filename = "Cell_cycle_Score_combine1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Signaling Scores compare all vs mettgf_tgf
```{r}
p <- VlnPlot(
  Fibro,
  features = c("Signaling_Score_all1", "Signaling_Score_mettgf_tgf1"),
  group.by = "condition",
  combine = TRUE
)
p
ggsave(
  filename = "Signaling_Score_combine1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Development Scores compare all vs mettgf_tgf
```{r}
p <- VlnPlot(
  Fibro,
  features = c("Development_Score_all1", "Development_Score_mettgf_tgf1"),
  group.by = "condition",
  combine = TRUE
)
p
ggsave(
  filename = "Development_Score_combine1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

## Visualizing Cell Death Scores compare all vs mettgf_tgf
```{r}
p <- VlnPlot(
  Fibro,
  features = c("Cell_death_Score_all1", "Cell_death_Score_mettgf_tgf1"),
  group.by = "condition",
  combine = TRUE
)
p
ggsave(
  filename = "Cell_death_Score_combine1.jpg",
  plot = p,
  width = 8,
  height = 6,
  dpi = 300
)
```

# Session information

```{r session-info, echo=FALSE}
sessioninfo::session_info()
```